---
title: Install R packages from Github through proxy
author: ''
date: '2017-08-12'
slug: install-r-packages-from-github-through-proxy
categories:
  - R
tags:
  - Tips in R
summary: "Guide to install R packages from Github through proxy"
---

The easiest way to install an R package is from R-Cran. However, as the number of R users is growing up very fast, many developers are contribute to R community through Github. And the standard way to install from Github is to use `devtools` package with `install_github()` command.

In many organization, especially financial organization like banking industry, sometimes it is very hard to install an R package from CRAN or Github due to security process. So, in this post I will share with you 2 ways to install an R package via proxy

##Method 1: Using `downloader` package

In this method, the only thing that you need to change is the Github repository of the R package needed to be installed.

For example, if you want to install package `easyRFM` from Github with the following link: `github.com/hoxo-m/easyRFM`, you can easily install wih the following codes:

```{r, eval = F}
#Step 1: Install downloader from CRAN
install.packages("downloader")
library(downloader)

#Step 2: Download R package from Github in format .tar.gz
download("https://github.com/hoxo-m/easyRFM/archive/master.tar.gz",
         "easyRFM.tar.gz")

#Step 3: Install from source
install.packages("easyRFM.tar.gz", repos = NULL, type = "source")
```

##Method 2: Using `httr` package to modify your proxy

In this method, you can modify your proxy using `httr` package. To modify your proxy, the key function is `set_config`.

```{r, eval = F} 
httr::set_config(use_proxy("your_proxy:your_port"))
```

For instance, if your proxy is `11.36.22.12` & your port is `8080`, you can modify your proxy as follows:

```{r, eval = F}
#Step 1: Call httr & devtools
library(httr)
library(devtools)

#Step 2: Modify proxy
set_config(use_proxy("11.36.22.12:8080"))

#Step 3: Install package directly from Github
install_github("rstats-db/odbc")

#Step 4: Reset your proxy if needed
reset_config()
```

Two methods above help me alot to install R packages, specially from Github. Hope that it will help you also to make your life easier to use R in your life of data scientists.
